---
title:         "Trends of SDR in Thessaloniki "
author:
  - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
  - Jane Doe^[Institution Two, jane@example.org]
abstract:
  "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."

documentclass:  article
classoption:    a4paper,oneside
fontsize:       12pt
geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
link-citations: yes
colorlinks:     yes
urlcolor:        blue

bibliography:    [references.bib]
biblio-style:    apalike

    
header-includes:
  - \usepackage{caption}
  - \usepackage{placeins}
  - \captionsetup{font=small}

output:
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         no
    latex_engine:     xelatex
    toc:              yes
    fig_width:        7
    fig_height:       4.5
  html_document:
    toc:        true
    fig_width:  7.5
    fig_height: 5
date: "`r format(Sys.time(), '%F')`"

---



```{r echo=F, include=F, cache=T}
knitr::opts_chunk$set(comment    = ""      )
knitr::opts_chunk$set(dev        = "pdf"   )
# knitr::opts_chunk$set(dev        = "png"    )
knitr::opts_chunk$set(out.width  = "80%"   )
knitr::opts_chunk$set(fig.align  = "center" )
knitr::opts_chunk$set(fig.pos    = '!h'    )

```


## Data description 

<!-- time span -->
Time span for CM21 is 1993-2022 and 2016-2022 for CHP1.

...... Measurements CHP1 CM21


<!-- screening -->
All measurements have been manually screened for inconsistencies and obviously 
problematic data have been excluded.
<!-- physical quantities -->
The conversions to irradiance was done using the calibration factor
.......chp1.... cmp21......
, and adjusting for the dark signal offset daily, using the nights 
before and after each day.
<!-- Quality assurance -->
A radiation data quality assurance procedure was applied adjusted for the site,
based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. 
Only data characterized with low quality indicators was rejected.
<!-- Data preparation -->
Due to the large measurement uncertainty near the horizon we exclude
all measurements with SZA greater than $85^\circ$. Moreover, due to 
nearby hills and buildings we remove data with azimuth angle between 
$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$, as at 
those cases the Sun can not viewed by the instruments.


## Data process

We take some steps in order to make the measurements comparable throughout
the dataset. 
<!-- 1au -->
We adjusted all measurements to the mean distance between Sun and Earth.
<!-- TSI -->
And, we make all measurements relative to the .........TSI.......... , 
using an homogenized timeseries of satellite TSI data...........

<!-- Clear Sky -->
We will use two dataset of our measurements, one for all sky conditions (ALL),
which is the whole available data, and one filtered for 'clear sky' 
conditions (CS). The latter was created, using a method adapted to our 
site as proposed by @Long2000 and by @Reno2016.




## Aggregation 

Both dataset were aggregate in SZA bins by am, pm, am+pm

by daily mean 

seasonal daily mean

Deseasonlized


## Longterm trends

```{r trackercomm, echo=F, fig.cap="tracker"}
knitr::include_graphics("./images/longtermtrendsALL-2.pdf")
```

This is it \@ref(fig:trackercomm).


\FloatBarrier


<!-- \tiny -->
```{r echo=F}
pp <- data.table::fread("./figures/tbl_longterm_trends.dat")
pander::pander(pp,
               caption = "Table of trends and stats")
```
<!-- \normalsize -->

## SZA trends

## Trends consistency


**END**


<!--chapter:end:Article.Rmd-->


```{r }
#!/usr/bin/env Rscript
```

---
title:         "Trends of SDR in Thessaloniki "
author:
  - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
  - Jane Doe^[Institution Two, jane@example.org]
abstract:
  "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."

documentclass:  article
classoption:    a4paper,oneside
fontsize:       10pt
geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
link-citations: yes
colorlinks:     yes

header-includes:
- \usepackage{caption}
- \usepackage{placeins}
- \captionsetup{font=small}

output:
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      no
    keep_tex:         no
    latex_engine:     xelatex
    toc:              yes
    fig_width:        7
    fig_height:       4.5
  html_document:
    toc:        true
    fig_width:  7.5
    fig_height: 5
date: "`r format(Sys.time(), '%F')`"

---



```{r echo=F, include=T}


####_  Document options _####

```
```{r echo=F, include=F}
knitr::opts_chunk$set(comment    = ""      )
knitr::opts_chunk$set(dev        = "pdf"   )
# knitr::opts_chunk$set(dev        = "png"    )
knitr::opts_chunk$set(out.width  = "80%"   )
knitr::opts_chunk$set(fig.align  = "center" )
knitr::opts_chunk$set(cache      =  T       )
# knitr::opts_chunk$set(fig.pos    = '!h'    )



```
```{r include=F, echo=F}
####  Set environment  ####
Sys.setenv(TZ = "UTC")
tic <- Sys.time()
Script.Name <- tryCatch({ funr::sys.script() },
                        error = function(e) { cat(paste("\nUnresolved script name: ", e),"\n\n")
                            return("Climatological_") })
if(!interactive()) {
    pdf(  file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".pdf", Script.Name))))
    sink( file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".out", Script.Name))), split=TRUE)
    filelock::lock(paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime//",  basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
}

par(pch = ".")


## FIXME this is for pdf output
# options(warn=-1) ## hide warnings
# options(warn=2)  ## stop on warnings

```
```{r echo=F, include=T}
####  External code  ####
library(data.table, quietly = T, warn.conflicts = F)
library(pander,     quietly = T, warn.conflicts = F)

## Functions from `https://github.com/thanasisn/IStillBreakStuff/tree/main/FUNCTIONS/R`
source("~/CODE/FUNCTIONS/R/sumNA.R")
source("~/CODE/FUNCTIONS/R/linear_regrassion_capture.R")
source("~/CODE/FUNCTIONS/R/trig_deg.R")


options(error = function() {
    if (interactive()) {
        # require("beepr"); beep(10)
        system("mplayer /usr/share/sounds/freedesktop/stereo/dialog-warning.oga", ignore.stdout = T, ignore.stderr = T)
        system("notify-send -u normal -t 30000 'R session' 'An error occured!'")
    }
})


####  Variables  ####
panderOptions('table.alignment.default', 'right')
panderOptions('table.split.table',        120   )

DATA_DIR  <- "/home/athan/DATA/Broad_Band"
CLEARdir  <- "/home/athan/DATA/Broad_Band/CS_id"
tag       <- paste0("Natsis Athanasios LAP AUTH ", strftime(Sys.time(), format = "%b %Y" ))
CS_file   <- "/home/athan/DATA/Common_application/Clear_Sky.Rds"

TEST      <- TRUE

col_DIR_att    <- "#2166ac"
col_DIR_transp <- "#9970ab"
col_GLB_att    <- "#1a9850"
pch_am         <-  1
pch_pm         <-  4
pch_ampm       <- 13 ## try 10
pch_daily      <- 19

MIN_ELEVA  <- 5  ## use data when sun above that
SZA_BIN    <- 1
MIN_N      <- 4
SEAS_MIN_N <- 3

SUM_THRES  <- 50
SZA_THRES  <- 3
```

## Data selection

```{r echo=F, include=T}

####  Get data from Clear sky id data  ####
input_files <- list.files( path    = CLEARdir,
                           pattern = "Clear_Sky_[0-9]{4}.Rds",
                           full.names = T )

if ( !file.exists(CS_file) | max(file.mtime(input_files)) > file.mtime(CS_file)) {
    cat(paste("Load data from Clear Sky proccess from original\n"))
    DATA <- data.table()
    for (af in input_files) {
        temp <- readRDS(af)
        ## drop some data
        temp$CHP1temp        <- NULL
        temp$CHP1tempSD      <- NULL
        temp$CHP1tempUNC     <- NULL
        temp$CS_ref          <- NULL
        temp$Pressure_Source <- NULL
        temp$chp1TempCF      <- NULL
        temp$wattDIR_tmp_cr  <- NULL
        temp$wattDIR_unc_NT  <- NULL
        temp$wattDIR_unc_WT  <- NULL
        temp$wattHOR_tmp_cr  <- NULL
        temp$wattHOR_unc_NT  <- NULL
        temp$wattHOR_unc_WT  <- NULL

        DATA <- rbind(temp, DATA, fill = TRUE)
        rm(temp)
    }
    ## this is used by old scripts
    myRtools::write_RDS(object = DATA, file = CS_file)
} else {
    cat(paste("Load data from Clear Sky proccess from parsed\n"))
    DATA <- readRDS(CS_file)
}
```

### Data range
Time data span `r range(DATA$Date)`

### Filter min elevation
Keep data with Sun elevation above `r MIN_ELEVA`

```{r }
DATA <- DATA[ Elevat >= MIN_ELEVA ]
```

### Bais paper obstacle filter

```{r }
DATA <- DATA[ ! (Azimuth > 35 & Azimuth < 120 & Elevat < 10) ]
```
```{r echo=F, include=T}
```

### Keep only data characterized as 'good' by the Radiation Quality control procedure

```{r echo=F, include=T}
DATA[ QCF_DIR != "good", wattDIR := NA ]
DATA[ QCF_DIR != "good", QCF_DIR := NA ]
DATA[ QCF_GLB != "good", wattGLB := NA ]
DATA[ QCF_GLB != "good", QCF_GLB := NA ]
DATA <- DATA[ ! (is.na(wattDIR) & is.na(wattGLB))  ]



## exclude some data for each instrument
# DATA[ QCF_DIR == "Extremely rare limits min (3)", wattDIR := NA ]
# DATA[ QCF_DIR == "Extremely rare limits min (3)", QCF_DIR := NA ]
#
# dirtb <- table(DATA$QCF_DIR)
# for (aty in names(dirtb)[!names(dirtb) %in% "good"]) {
#     temp <- DATA[ QCF_DIR == aty ]
#     cat(nrow(temp),aty,"\n")
# }
# glbtb <- table(DATA$QCF_GLB)
# for (aty in names(glbtb)[!names(glbtb) %in% "good"]) {
#     temp <- DATA[ QCF_GLB == aty ]
#     cat(nrow(temp),aty,"\n")
# }
```

## Data preparation
### Move measurements to mean earth distance

```{r }
DATA[ , wattDIR_1au := wattDIR * (sun_dist ^ 2) ]
DATA[ , wattGLB_1au := wattGLB * (sun_dist ^ 2) ]
```
```{r echo=F, include=T}
```

### Relative to actual TSI at 1au variable representation

```{r }
DATA[ , DIR_att := wattDIR_1au / tsi_1au_comb ]
DATA[ , GLB_att := wattGLB_1au / tsi_1au_comb ]
```

Using the actual values gives similar trends.

```{r echo=F, include=T}

# #' ### Use original variable representation
# DATA[ , DIR_att := wattDIR_1au ]
# DATA[ , GLB_att := wattGLB_1au ]
# #+ echo=F, include=T




# #' ### Ground effect removal?
# #' Aerosol direct effects on global solar shortwave irradiance at high mountainous station Musala Bulgaria_Nojarov2021.pdf
# DATA$wattGLB_1au <- DATA$wattGLB_1au / cosde(DATA$SZA)
# DATA$wattDIR_1au <- DATA$wattDIR_1au / cosde(DATA$SZA)
# #+ echo=F, include=T
```

### Calculate Bouguer atmospheric transparency
see: Changes in solar radiation and their influence on temperature trend in Estonia 1955 2007_Russak2009.pdf

```{r }
DATA[, DIR_transp := ( wattDIR_1au / tsi_1au_comb ) ^ ( 1 / cosde(SZA) ) ]
```
```{r echo=F, include=T}



## fix noon just in case
DATA[ Azimuth <= 180 , preNoon := TRUE  ]
DATA[ Azimuth >  180 , preNoon := FALSE ]
```

### Split data to Clear Sky and non Clear sky
Method based and adapted to site from: Identification of Periods of Clear Sky
Irradiance in Time Series of GHI Measurements
_Matthew J. Reno and Clifford W. Hansen_.

```{r echo=F, include=T}
DATA_all   <- DATA

wecare     <- grep("CSflag_", names(DATA), value = T)
## use only cm21 flags
wecare     <- grep("_11", wecare, invert = T, value = T)
DATA_Clear <- DATA[ rowSums( DATA[, ..wecare ], na.rm = T ) == 0, ]
```
```{r echo=F, include=T}

## old flags
# DATA_Clear <- DATA_all[ CSflag == 0 ]
rm(DATA)



# #### . . My CSid method   ####
# #'
# #' | CS Flag | Test |
# #' |:-------:|:--------------------------------------------------------------|
# #' |   NA    | Undefined, untested                                           |
# #' |    0    | Passed as clear sky                                           |
# #' |    1    | Mean value of irradiance during the time period (MeanVIP)     |
# #' |    2    | Max Value of Irradiance during the time Period (MaxVIP)       |
# #' |    3    | Variability in irradiance by the length (VIL)                 |
# #' |    4    | Variance of Changes in the Time series (VCT)                  |
# #' |    5    | Variability in the Shape of the irradiance Measurements (VSM) |
# #' |    6    | Low Direct Irradiance limit (LDI)                             |
# #' |    7    | Low Global Irradiance limit (LGI)                             |
# #' |    8    | Too Few CS point for the day (FCS)                            |
# #' |    9    | Too Few data points for the day                               |
# #' |   10    | Missing Data                                                  |
# #' |   11    | Direct irradiance simple threshold                            |






#### 2. Long term by SZA ####
```

## 2. Long term by SZA

```{r }
##TODO t-test and significance
```

#### Calculate daily SZA means ####

```{r echo=F, include=T}

ALL_daily_mean <- DATA_all[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                               GLB_att       = mean(GLB_att,    na.rm = T),
                               DIR_transp    = mean(DIR_transp, na.rm = T),
                               DIR_att_sd    = sd(  DIR_att,    na.rm = T),
                               GLB_att_sd    = sd(  GLB_att,    na.rm = T),
                               DIR_transp_sd = sd(DIR_transp, na.rm = T),
                               doy           = yday(Date),
                               GLB_att_N     = sum(!is.na(GLB_att)),
                               DIR_att_N     = sum(!is.na(DIR_att))  ),
                           by = .(SZA     = (SZA - SZA_BIN / 2 ) %/% SZA_BIN,
                                  Date    = Day,
                                  preNoon = preNoon  ) ]


CLEAR_daily_mean <- DATA_Clear[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                                   GLB_att       = mean(GLB_att,    na.rm = T),
                                   DIR_transp    = mean(DIR_transp, na.rm = T),
                                   DIR_att_sd    = sd(  DIR_att,    na.rm = T),
                                   GLB_att_sd    = sd(  GLB_att,    na.rm = T),
                                   DIR_transp_sd = sd(DIR_transp, na.rm = T),
                                   doy           = yday(Date),
                                   GLB_att_N     = sum(!is.na(GLB_att)),
                                   DIR_att_N     = sum(!is.na(DIR_att))  ),
                               by = .(SZA     = (SZA - SZA_BIN / 2 ) %/% SZA_BIN,
                                      Date    = Day,
                                      preNoon = preNoon ) ]
```

#### Margin of error calculation for 0.95 confidence interval ####

```{r echo=T, include=T}
CONF_INTERV <- .95
conf_param  <- 1-(1-CONF_INTERV)/2
suppressWarnings({
ALL_daily_mean[,   DIR_att_EM   := qt(conf_param,df=DIR_att_N -1) * DIR_att_sd    / sqrt(DIR_att_N)]
ALL_daily_mean[,   GLB_att_EM   := qt(conf_param,df=GLB_att_N -1) * GLB_att_sd    / sqrt(GLB_att_N)]
ALL_daily_mean[,   DIR_transp_EM:= qt(conf_param,df=DIR_att_N -1) * DIR_transp_sd / sqrt(DIR_att_N)]
CLEAR_daily_mean[, DIR_att_EM   := qt(conf_param,df=DIR_att_N -1) * DIR_att_sd    / sqrt(DIR_att_N)]
CLEAR_daily_mean[, GLB_att_EM   := qt(conf_param,df=GLB_att_N -1) * GLB_att_sd    / sqrt(GLB_att_N)]
CLEAR_daily_mean[, DIR_transp_EM:= qt(conf_param,df=DIR_att_N -1) * DIR_transp_sd / sqrt(DIR_att_N)]
})
```
```{r echo=F, include=F}
```

#### Exclude means with less than `r SZA_THRES` data points

```{r echo=F, include=T}
ALL_daily_mean[   DIR_att_N <= SZA_THRES, DIR_att    := NA ]
ALL_daily_mean[   GLB_att_N <= SZA_THRES, GLB_att    := NA ]
ALL_daily_mean[   DIR_att_N <= SZA_THRES, DIR_transp := NA ]
CLEAR_daily_mean[ DIR_att_N <= SZA_THRES, DIR_att    := NA ]
CLEAR_daily_mean[ GLB_att_N <= SZA_THRES, GLB_att    := NA ]
CLEAR_daily_mean[ DIR_att_N <= SZA_THRES, DIR_transp := NA ]
```

#### Calculate daily seasonal values by SZA ####

```{r echo=F, include=T}

ALL_daily_seas <-
    ALL_daily_mean[, .(DIR_att_seas    = mean(DIR_att,    na.rm = T),
                       GLB_att_seas    = mean(GLB_att,    na.rm = T),
                       DIR_transp_seas = mean(DIR_transp, na.rm = T),
                       DIR_att_sd_seas = sd(  DIR_att,    na.rm = T),
                       GLB_att_sd_seas = sd(  GLB_att,    na.rm = T),
                       GLB_att_N_seas  = sum(!is.na(GLB_att)),
                       DIR_att_N_seas  = sum(!is.na(DIR_att))  ),
                   by = .( doy, SZA, preNoon ) ]

CLEAR_daily_seas <-
    CLEAR_daily_mean[, .(DIR_att_seas    = mean(DIR_att,    na.rm = T),
                         GLB_att_seas    = mean(GLB_att,    na.rm = T),
                         DIR_transp_seas = mean(DIR_transp, na.rm = T),
                         DIR_att_sd_seas = sd(  DIR_att,    na.rm = T),
                         GLB_att_sd_seas = sd(  GLB_att,    na.rm = T),
                         GLB_att_N_seas  = sum(!is.na(GLB_att)),
                         DIR_att_N_seas  = sum(!is.na(DIR_att))  ),
                     by = .( doy, SZA, preNoon ) ]



```
```{r echo=F, include=F}
## ~ Plots longterm  ####
plot( ALL_daily_mean$Date, ALL_daily_mean$DIR_att )
plot( ALL_daily_mean$Date, ALL_daily_mean$GLB_att )
plot( ALL_daily_mean$Date, ALL_daily_mean$DIR_transp )

plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_att )
plot( ALL_daily_mean$doy, ALL_daily_mean$GLB_att )
plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_transp )

plot( ALL_daily_mean$doy, ALL_daily_mean$GLB_att_N)
plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_att_N)

hist(ALL_daily_mean$DIR_att_N)
hist(ALL_daily_mean$GLB_att_N)


plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$DIR_att )
plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$GLB_att )
plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$DIR_transp )

plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_att )
plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$GLB_att )
plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_transp )

plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$GLB_att_N)
plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_att_N)

hist( CLEAR_daily_mean$DIR_att_N)
hist( CLEAR_daily_mean$GLB_att_N)

## ~ Plots seasonal ####

plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_att_seas )
plot( ALL_daily_seas$doy, ALL_daily_seas$GLB_att_seas )
plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_transp_seas )
plot( ALL_daily_seas$doy, ALL_daily_seas$GLB_att_N_seas)
plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_att_N_seas)

plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_att_seas )
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$GLB_att_seas )
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_transp_seas )
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$GLB_att_N_seas)
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_att_N_seas)
```

#### Calculate seasonal anomaly ####

```{r echo=F, include=F}

ALL_daily_seas   <- merge(  ALL_daily_mean, ALL_daily_seas,   by = c("doy", "SZA", "preNoon"),  all = T)
CLEAR_daily_seas <- merge(CLEAR_daily_mean, CLEAR_daily_seas, by = c("doy", "SZA", "preNoon"), all = T)

setorder(ALL_daily_seas,Date)
setorder(CLEAR_daily_seas,Date)


## anomaly
# ALL_daily_seas[   , DIR_att    := DIR_att    - DIR_att_seas    ]
# ALL_daily_seas[   , GLB_att    := GLB_att    - GLB_att_seas    ]
# ALL_daily_seas[   , DIR_transp := DIR_transp - DIR_transp_seas ]
# CLEAR_daily_seas[ , DIR_att    := DIR_att    - DIR_att_seas    ]
# CLEAR_daily_seas[ , GLB_att    := GLB_att    - GLB_att_seas    ]
# CLEAR_daily_seas[ , DIR_transp := DIR_transp - DIR_transp_seas ]

## relative anomaly
```
```{r echo=T, include=T}
ALL_daily_seas[  , DIR_att   := 100 * ( DIR_att    - DIR_att_seas    ) / DIR_att_seas    ]
ALL_daily_seas[  , GLB_att   := 100 * ( GLB_att    - GLB_att_seas    ) / GLB_att_seas    ]
ALL_daily_seas[  , DIR_transp:= 100 * ( DIR_transp - DIR_transp_seas ) / DIR_transp_seas ]
CLEAR_daily_seas[, DIR_att   := 100 * ( DIR_att    - DIR_att_seas    ) / DIR_att_seas    ]
CLEAR_daily_seas[, GLB_att   := 100 * ( GLB_att    - GLB_att_seas    ) / GLB_att_seas    ]
CLEAR_daily_seas[, DIR_transp:= 100 * ( DIR_transp - DIR_transp_seas ) / DIR_transp_seas ]
```
```{r echo=F, include=F}



####  Plot of SZA trends for all year ####
```

\newpage
## Plot of SZA trends

```{r echo=F, include=F}
timefactor <- 1
vars <- c("DIR_att", "GLB_att", "DIR_transp")
dbs  <- c("ALL_daily_seas", "CLEAR_daily_seas")

gather <- data.frame()

for (DBn in dbs) {
    DB <- get(DBn)
    for (avar in vars) {
        for (anoon in unique( DB$preNoon)) {
            for (asza in unique( DB$SZA )) {

                dataset <- DB[ SZA == asza & preNoon == anoon ]

                if (sum(!is.na(dataset[[avar]])) <= 1) next()

                lm1 <- lm( dataset[[avar]] ~ dataset$Date )

                gather <- rbind(gather,
                                data.frame(
                                    linear_regression_capture(lm1),
                                    preNoon   = anoon,
                                    SZA       = asza,
                                    DATA      = DBn,
                                    var       = avar,
                                    N         = sum(!is.na(dataset[[avar]]))
                                ))

                # plot(dataset$Date, dataset[[avar]], pch  = "." )
                # abline(lm1)
                # fit <- lm1[[1]]
                # legend('top', lty = 1, bty = "n",
                #        paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
                # title(paste(DBn, asza, anoon))
            }
        }
    }
}

```
```{r echo=F, include=F}
hist(gather$N[gather$N>50], breaks = 100)

szatrends <- gather

szatrends <- data.table(szatrends)
setorder(szatrends,SZA)

szatrends[ var == "DIR_att",    col := col_DIR_att    ]
szatrends[ var == "GLB_att",    col := col_GLB_att    ]
szatrends[ var == "DIR_transp", col := col_DIR_transp ]
szatrends[ preNoon == T, pch := pch_am ]
szatrends[ preNoon == F, pch := pch_pm ]



hist(szatrends[DATA==dbs[1],N], breaks = 100)
hist(szatrends[DATA==dbs[2],N], breaks = 100)

hist(szatrends[var==vars[1],N], breaks = 100)
hist(szatrends[var==vars[2],N], breaks = 100)

# szatrends <- szatrends[ N > 50]


plot(szatrends$SZA,szatrends$N)

test1 <- szatrends[ DATA == "CLEAR_daily_seas" & var == "DIR_att" ]
test2 <- szatrends[ DATA == "CLEAR_daily_seas" & var == "GLB_att" ]
plot(test1$SZA, test1$N, pch =19)
abline(h=50)
plot(test2$SZA, test2$N, pch =19)
abline(h=300)

# szatrends[ var == "GLB_att"    & N <= 300, slope := NA ]
# szatrends[ var == "DIR_att"    & N <=  50, slope := NA ]
# szatrends[ var == "DIR_transp" & N <=  50, slope := NA ]





## stats vars to plot
wecare <- grep( "^slope|^N",names(szatrends),ignore.case = T, value = T)


```
```{r szatrends, echo=F, include=T}
## ALL - CS
for (type in unique(szatrends$DATA)) {
    ## DIR - GLB - transp
    for (avar in unique(szatrends$var)) {
        ## statistic variable
        for (awe in wecare) {
            awename <- gsub("(\\D)(\\D+)", "\\U\\1\\L\\2", sub("\\."," ", awe), perl = TRUE)

            ## select All/CS and DIR/GLB/trans
            subdata <- szatrends[ szatrends$DATA == type &    ##
                                  szatrends$var  == avar , ]

            xlim <- range( subdata$SZA,    na.rm = T )
            ylim <- range( subdata[[awe]], na.rm = T )

            pam  <- subdata[ preNoon == T ]
            ppm  <- subdata[ preNoon == F ]

            plot(1, type="n", xlab="SZA", ylab=awename, xlim=xlim, ylim=ylim )
            abline(h=0)
            title(paste(awename,type, avar),cex=0.9)

            # lines(pam$SZA, pam[[awe]], pch =  pch_am, col = pam$col, type = "b")
            # lines(pam$SZA, ppm[[awe]], pch =  pch_pm, col = pam$col, type = "b")

            lines(pam$SZA, pam[[awe]], pch =  pch_am, col = 2, type = "b")
            lines(ppm$SZA, ppm[[awe]], pch =  pch_pm, col = 3, type = "b")

            legend("top",
                   legend = c("Morning",       "Evening"),
                   # col    = c(unique(pam$col), unique(ppm$col)),
                   col    = c(2, 3),
                   pch    = c(unique(pam$pch), unique(ppm$pch)), ncol = 2, bty = "n")
        }

#
#         subdata <- szatrends[ szatrends$DATA == type & szatrends$var == avar , ]
#
#         plot(subdata$SZA, subdata$slope, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$slope.sd, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope sd",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$slope.t, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope t",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$slope.p, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope p",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$Rsqrd, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope Rsq",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$N, col = subdata$preNoon, pch = 19)
#         title(paste("N",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")

    }
}








####  Plot of SZA trends for each season of year ####
```

\newpage
## Plot of SZA trends for each season of year

```{r echo=F, include=F}
timefactor  <- 1  ## to display % per year
vars        <- c("DIR_att", "GLB_att", "DIR_transp")
dbs         <- c("ALL_daily_seas", "CLEAR_daily_seas")
season      <- c("Winter", "Spring", "Summer", "Automn")
gather_seas <- data.frame()

for (DBn in dbs) {
    DB <- get(DBn)

    ## set seasons in each data base
    DB[ month(Date) %in% c(12, 1, 2), Season := "Winter"]
    DB[ month(Date) %in% c( 2, 4, 5), Season := "Spring"]
    DB[ month(Date) %in% c( 6, 7, 8), Season := "Summer"]
    DB[ month(Date) %in% c( 9,10,11), Season := "Automn"]

    for (ase in season) {
        for (avar in vars) {
            for (anoon in unique( DB$preNoon)) {
                for (asza in unique( DB$SZA )) {

                    dataset <- DB[ SZA == asza & preNoon == anoon & season == ase ]

                    if (sum(!is.na(dataset[[avar]])) <= 1) next()

                    lm1 <- lm( dataset[[avar]] ~ dataset$Date )

                    gather_seas <- rbind(gather_seas,
                                    data.frame(
                                        linear_regression_capture(lm1),
                                        preNoon   = anoon,
                                        SZA       = asza,
                                        DATA      = DBn,
                                        var       = avar,
                                        Season    = ase,
                                        N         = sum(!is.na(dataset[[avar]]))
                                    ))

                    # plot(dataset$Date, dataset[[avar]], pch  = "." )
                    # abline(lm1)
                    # fit <- lm1[[1]]
                    # legend('top', lty = 1, bty = "n",
                    #        paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
                    # title(paste(DBn, asza, anoon))
                }
            }
        }
    }
}

```
```{r echo=F, include=F}
hist(gather_seas$N[gather_seas$N>50], breaks = 100)

szatrends_seas <- data.table(gather_seas)
setorder(szatrends_seas,SZA)


## define plot colors
szatrends_seas[ var == "DIR_att",    col := col_DIR_att    ]
szatrends_seas[ var == "GLB_att",    col := col_GLB_att    ]
szatrends_seas[ var == "DIR_transp", col := col_DIR_transp ]
szatrends_seas[ preNoon == T, pch := pch_am ]
szatrends_seas[ preNoon == F, pch := pch_pm ]





hist(szatrends_seas[DATA==dbs[1],N], breaks = 100)
hist(szatrends_seas[DATA==dbs[2],N], breaks = 100)
hist(szatrends_seas[var==vars[1],N], breaks = 100)
hist(szatrends_seas[var==vars[2],N], breaks = 100)

# szatrends_seas <- szatrends_seas[ N > 50]

plot(szatrends_seas$SZA,szatrends_seas$N)

test <- szatrends_seas[ DATA == "CLEAR_daily_seas" & var == "DIR_att" ]
plot(test$SZA, test$N, pch =19)
abline(h=50/4)

szatrends[ N <= 30, slope := NA]


test1 <- szatrends_seas[ DATA == "CLEAR_daily_seas" & var == "DIR_att" ]
test2 <- szatrends_seas[ DATA == "CLEAR_daily_seas" & var == "GLB_att" ]
plot(test1$SZA, test1$N, pch =19)
abline(h=50/4)
plot(test2$SZA, test2$N, pch =19)
abline(h=300/4)

# szatrends[ var == "GLB_att"    & N <= 300, slope := NA ]
# szatrends[ var == "DIR_att"    & N <=  50, slope := NA ]
# szatrends[ var == "DIR_transp" & N <=  50, slope := NA ]



## stats vars to plot
wecare <- grep( "^slope|^N",names(szatrends_seas),ignore.case = T, value = T)


```
```{r szatrendsseas, echo=F, include=T}
## Winter - Summer ....
for (ase in season) {
    ## ALL - Clear sky
    for (type in unique(szatrends_seas$DATA)) {
        ## DIR - GLB - transp
        for (avar in unique(szatrends_seas$var)) {

            ## statistic variable
            for (awe in wecare) {
                awename <- gsub("(\\D)(\\D+)", "\\U\\1\\L\\2", sub("\\."," ", awe), perl = TRUE)

                ## select All/CS  DIR/GLB/trans winter/summer
                subdata <- szatrends_seas[ DATA   == type &
                                           var    == avar &
                                           Season == ase    , ]

                xlim <- range( subdata$SZA,    na.rm = T )
                ylim <- range( subdata[[awe]], na.rm = T )

                pam  <- subdata[ preNoon == T ]
                ppm  <- subdata[ preNoon == F ]

                plot(1, type="n", xlab="SZA", ylab=awename, xlim=xlim, ylim=ylim )
                abline(h=0)
                title(paste(ase,awename,type, avar),cex=0.9)

                # lines(pam$SZA, pam[[awe]], pch =  pch_am, col = pam$col, type = "b")
                # lines(pam$SZA, ppm[[awe]], pch =  pch_pm, col = pam$col, type = "b")

                lines(pam$SZA, pam[[awe]], pch =  pch_am, col = 2, type = "b")
                lines(ppm$SZA, ppm[[awe]], pch =  pch_pm, col = 3, type = "b")

                legend("top",
                       legend = c("Morning",       "Evening"),
                       # col    = c(unique(pam$col), unique(ppm$col)),
                       col    = c(2, 3),
                       pch    = c(unique(pam$pch), unique(ppm$pch)), ncol = 2, bty = "n")
            }



#
#             subdata <- szatrends_seas[ DATA   == type &
#                                        var    == avar &
#                                        Season == ase    , ]
#
#
#             pam <- subdata[ preNoon == T ]
#             ppm <- subdata[ preNoon == F ]
#
#             plot(subdata$SZA, subdata$slope, col = subdata$col, pch = subdata$pch)
#             abline(h=0)
#             title(paste(ase, "Slope",type, avar),cex=0.9)
#             legend("top",
#                    legend = c("Morning", "Evening"),
#                    col    = c(2 , 3),
#                    pch    = c(pch_am,     pch_pm), ncol = 2, bty = "n")
#
            # plot(subdata$SZA, subdata$slope.sd, col = subdata$preNoon, pch = 19)
            # abline(h=0)
            # title(paste("Slope sd",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")
            #
            # plot(subdata$SZA, subdata$slope.t, col = subdata$preNoon, pch = 19)
            # abline(h=0)
            # title(paste("Slope t",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")
            #
            # plot(subdata$SZA, subdata$slope.p, col = subdata$preNoon, pch = 19)
            # abline(h=0)
            # title(paste("Slope p",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")
            #
            # plot(subdata$SZA, subdata$Rsqrd, col = subdata$preNoon, pch = 19)
            # abline(h=0)
            # title(paste("Slope Rsq",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")
            #
            # plot(subdata$SZA, subdata$N, col = subdata$preNoon, pch = 19)
            # title(paste("N",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")

        }
    }

}
```

**END**

```{r include=T, echo=F}
tac <- Sys.time()
cat(sprintf("%s %s@%s %s %f mins\n\n",Sys.time(),Sys.info()["login"],Sys.info()["nodename"],Script.Name,difftime(tac,tic,units="mins")))
```



---
title: DHI_GHI_sza_trends.R
author: athan
date: '2022-10-07'

---

<!--chapter:end:DHI_GHI_sza_trends.spin.Rmd-->

