---
title:         "Trends of SDR in Thessaloniki "
author:
  - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
  - Jane Doe^[Institution Two, jane@example.org]
abstract:
  "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."

documentclass:  article
classoption:    a4paper,oneside
fontsize:       12pt
geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
link-citations: yes
colorlinks:     yes
urlcolor:        blue

bibliography:    [references.bib]
biblio-style:    apalike

    
header-includes:
  - \usepackage{caption}
  - \usepackage{placeins}
  - \captionsetup{font=small}

output:
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         no
    latex_engine:     xelatex
    toc:              yes
    fig_width:        7
    fig_height:       4.5
  html_document:
    toc:        true
    fig_width:  7.5
    fig_height: 5
date: "`r format(Sys.time(), '%F')`"

---



```{r echo=F, include=F, cache=T}
knitr::opts_chunk$set(comment    = ""      )
knitr::opts_chunk$set(dev        = "pdf"   )
# knitr::opts_chunk$set(dev        = "png"    )
knitr::opts_chunk$set(out.width  = "80%"   )
knitr::opts_chunk$set(fig.align  = "center" )
knitr::opts_chunk$set(fig.pos    = '!h'    )

```


## Data description 

<!-- time span -->
Time span for CM21 is 1993-2022 and 2016-2022 for CHP1.

...... Measurements CHP1 CM21


<!-- screening -->
All measurements have been manually screened for inconsistencies and obviously 
problematic data have been excluded.
<!-- physical quantities -->
The conversions to irradiance was done using the calibration factor
.......chp1.... cmp21......
, and adjusting for the dark signal offset daily, using the nights 
before and after each day.
<!-- Quality assurance -->
A radiation data quality assurance procedure was applied adjusted for the site,
based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. 
Only data characterized with low quality indicators was rejected.
<!-- Data preparation -->
Due to the large measurement uncertainty near the horizon we exclude
all measurements with SZA greater than $85^\circ$. Moreover, due to 
nearby hills and buildings we remove data with azimuth angle between 
$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$, as at 
those cases the Sun can not viewed by the instruments.


## Data process

We take some steps in order to make the measurements comparable throughout
the dataset. 
<!-- 1au -->
We adjusted all measurements to the mean distance between Sun and Earth.
<!-- TSI -->
And, we make all measurements relative to the .........TSI.......... , 
using an homogenized timeseries of satellite TSI data...........

<!-- Clear Sky -->
We will use two dataset of our measurements, one for all sky conditions (ALL),
which is the whole available data, and one filtered for 'clear sky' 
conditions (CS). The latter was created, using a method adapted to our 
site as proposed by @Long2000 and by @Reno2016.




## Aggregation 

Both dataset were aggregate in SZA bins by am, pm, am+pm

by daily mean 

seasonal daily mean

Deseasonlized


## Longterm trends

```{r trackercomm, echo=F, fig.cap="tracker"}
knitr::include_graphics("./images/longtermtrendsALL-2.pdf")
```

This is it \@ref(fig:trackercomm).


\FloatBarrier


<!-- \tiny -->
```{r echo=F}
pp <- data.table::fread("./figures/tbl_longterm_trends.dat")
pander::pander(pp,
               caption = "Table of trends and stats")
```
<!-- \normalsize -->

## SZA trends

## Trends consistency


**END**

