---
title:         "Trends of SDR in Thessaloniki "
author:
  - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
  - Jane Doe^[Institution Two, jane@example.org]
abstract:
  "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."

documentclass:  article
classoption:    a4paper,oneside
fontsize:       12pt
geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
link-citations: yes
colorlinks:     yes
urlcolor:        blue

bibliography:    [references.bib]
biblio-style:    apalike

    
header-includes:
  - \usepackage{caption}
  - \usepackage{placeins}
  - \captionsetup{font=small}

output:
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         no
    latex_engine:     xelatex
    toc:              yes
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
  html_document:
    toc:        true
    fig_width:  7.5
    fig_height: 5
date: "`r format(Sys.time(), '%F')`"

---



```{r echo=F, include=F, cache=T}
knitr::opts_chunk$set(comment    = ""      )
knitr::opts_chunk$set(dev        = "pdf"   )
# knitr::opts_chunk$set(dev        = "png"    )
knitr::opts_chunk$set(out.width  = "80%"   )
knitr::opts_chunk$set(fig.align  = "center" )
knitr::opts_chunk$set(fig.pos    = '!h'    )

```


```{r echo=F, include=T}
## variables we use
source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
```




# Intoduction.

<!-- time span -->
Time span for CM21 is 1993-2022 and 2016-2022 for CHP1.

goal

basic consepts



# Data and methodology.

## Preperation.

<!-- measurement -->
For both CM21 and CHP1 instruments, the mean signal and the standard deviation over one minute is recorded every one minute.
<!-- screening -->
A manual screening processes, removes inconsistencies and erroneous recordings, that can occur on a long contentious operation, like this one.
<!-- physical quantities -->
The conversion to irradiance was done using the available calibrations factors. For CHP1, there is only the initial one by the manufacture, and for CM21 there are seven calibration procedure in the span of the period we investigate.
<!-- Dark signal -->
We have adjusted the measurements for the dark signal offset, by a daily value, using the nights before and after each day.
<!-- Quality assurance -->
A radiation data quality assurance procedure was applied, adjusted for the site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. Only data characterized as good quality were used in this stud.
<!-- Data preparation -->
Due to larger measurement uncertainty near the horizon, we exclude
all measurements with SZA greater than $85^\circ$. 
Moreover, due to some obstructions around the site (hills and buildings), we remove data with Azimuth angle between 
$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$. At 
those cases the Sun, were not visible by the instruments.


## Data process.

To make the measurements comparable throughout the dataset. 
<!-- 1au -->
We adjusted all 1-minute values to the mean Sun and Earth distance.
<!-- TSI -->
<!-- And, we made all measurements relative to the TSI at 1au, in order to compensate for the the Suns intensity variability, using an homogenized timeseries of satellite TSI observations. As a result, we processed all radiation data as a fraction of Suns TSI. -->


<!-- data selection -->
We select all data with $\text{SZA} < `r 90-MIN_ELEVA`^\circ$, in order to avoid measurements near the horizon. Where the signal is very low and the disturbances more frequent.
<!-- avoid obstacles -->
Similar, we exclude values data with Azimuth angle between 
$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$. Where, due to some obstructions around the site (hills and buildings) the Sun, were not visible by the instruments. . At 
those cases.



<!-- Clear Sky -->
In this analysis, we need .............
, we using two selection of our measurements dataset, one with all sky conditions (ALL),
consisting of the whole available data, and one filtered for 'clear sky' 
conditions (CS). The latter was created, using a method adapted to our 
site, and was proposed by @Long2000 and by @Reno2016.




# Results

## Long term trends

## Zenith angle dependance and years seasons.

## Consistency of the trends





# Conclusions



---------------------------




## Aggregation 

Both dataset were aggregate in $1^\circ$ SZA bins by am, pm, am+pm

some data was rejected.......

by daily mean 

daily sum

seasonal daily mean

<!-- Deseasonalization -->
We remove the natural occurring seasonal variation, by calculating the mean
values for the corresponding time step, and subtracting them .........



## Long-term trends

```{r trackercomm, echo=F, fig.cap="tracker"}
knitr::include_graphics("./images/longtermtrendsALL-2.pdf")
```

This is it \@ref(fig:trackercomm).


\FloatBarrier


<!-- \tiny -->
```{r echo=F}
pp <- data.table::fread("./figures/tbl_longterm_trends.dat")
pander::pander(pp,
               caption = "Table of trends and stats")
```
<!-- \normalsize -->


**END**

