---
title: Full title of the paper (Capitalized)
author:
  - name:  Athanasios Natsis
    affil: 1,* # \ddagger
    orcid: 0000-0002-5199-4119
  - name: Alkiviadis Bais
    affil: 2, #\dagger, # \ddagger
affiliation:
  - num: 1
    address: |
      Aristotle University of Thessaloniki - 
      Laboratory of Atmospheric Physics,
      Campus Box 149, 54124 Thessaloniki, Greece
    email: natsisphysicist@gmail.com
  - num: 2
    address: |
      Your department
      Street, City, Country
#     email: abais@auth.gr
# author citation list in chicago format
authorcitation: |
  Natsis, A.; Bais, A.
# firstnote to eighthnote
# firstnote: |
#   Current address: Updated affiliation
# secondnote: |
#   These authors contributed equally to this work.
correspondence: |
  natsisphysicist@gmail.com;
#   Tel.: +XX-000-00-0000.
# document options
journal: applsci
type:    article
status:  submit
# front matter
simplesummary: |
  A Simple summary goes here.
abstract: |
  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
  While monitoring the long term variability, there are observations of upward and downward SDR trends on different locations around the world for different time periods.
  Periods of positive tredns are refered as brightening periods and with negative tredns as dimming periods.
  We studied 29 years of CHP1 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
  We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trends, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
  We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
# back matter
keywords: |
  GHI; SDR; Solar Brigthening/Dimming (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
funding: |
  Please add: ``This research received no external funding'' or ``This research 
  was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded 
  by XXX''. Check carefully that the details given are accurate and use the 
  standard spelling of funding agency names at 
  \url{https://search.crossref.org/funding}, any errors may affect your future 
  funding.
institutionalreview: |
  In this section, you should add the Institutional Review Board Statement and 
  approval number, if relevant to your study. You might choose to exclude 
  this statement if the study did not require ethical approval. Please note 
  that the Editorial Office might ask you for further information. Please add 
  “The study was conducted in accordance with the Declaration of Helsinki, 
  and approved by the Institutional Review Board (or Ethics Committee) of 
  NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies 
  involving humans. OR “The animal study protocol was approved by the 
  Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE 
  (protocol code XXX and date of approval).” for studies involving animals. 
  OR “Ethical review and approval were waived for this study due to REASON 
  (please provide a detailed justification).” OR “Not applicable” for
   studies not involving humans or animals.
# informedconsent: |
#   Any research article describing a study involving humans should contain this 
#   statement. Please add ``Informed consent was obtained from all subjects 
#   involved in the study.'' OR ``Patient consent was waived due to REASON 
#   (please provide a detailed justification).'' OR ``Not applicable'' for 
#   studies not involving humans. You might also choose to exclude this statement 
#   if the study did not involve humans.
#   
#   Written informed consent for publication must be obtained from participating 
#   patients who can be identified (including by the patients themselves). Please 
#   state ``Written informed consent has been obtained from the patient(s) to 
#   publish this paper'' if applicable.
dataavailability: |
  We encourage all authors of articles published in MDPI journals to share 
  their research data. In this section, please provide details regarding where 
  data supporting reported results can be found, including links to publicly 
  archived datasets analyzed or generated during the study. Where no new data 
  were created, or where data is unavailable due to privacy or ethical 
  re-strictions, a statement is still required. Suggested Data Availability 
  Statements are available in section “MDPI Research Data Policies” at 
  \url{https://www.mdpi.com/ethics}.
conflictsofinterest: |
  Declare conflicts of interest or state 'The authors declare no conflict of 
  interest.' Authors must identify and declare any personal circumstances or
  interest that may be perceived as inappropriately influencing the
  representation or interpretation of reported research results. Any role of the
  funding sponsors in the design of the study; in the collection, analyses or 
  interpretation of data in the writing of the manuscript, or in the decision to 
  publish the results must be declared in this section. If there is no role, 
  please state 'The founding sponsors had no role in the design of the study; 
  in the collection, analyses, or interpretation of data; in the writing of the 
  manuscript, an in the decision to publish the results'.
# sampleavailability: |
#   Samples of the compounds ...... are available from the authors.
supplementary: |
 The following supporting information can be downloaded at:  
 \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: SDR 
    long: Solar downward radiation
bibliography: manualreferences.bib
appendix:     appendix.tex
endnotes:     false
output: 
  rticles::mdpi_article:
    extra_dependencies:  longtable 
    keep_tex:            yes
    
---



```{r echo=F, include=T, warning=FALSE, message=FALSE}
suppressMessages(
    res <- rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
)
## document configurations 
knitr::opts_chunk$set(comment    = ""      )
# knitr::opts_chunk$set(dev        = "pdf"   )
knitr::opts_chunk$set(out.width  = "70%"   )
knitr::opts_chunk$set(fig.align  = "center")
knitr::opts_chunk$set(fig.pos    = 'h!'    )
options(knitr.graphics.auto_pdf = FALSE)
library(pander)
library(data.table)
library(kableExtra)
library(dplyr)
## load data from trends analysis
load("./data/common_data_14_2.Rda")
load("./data/Input_1_longterm_trends.Rda")
## number of measurements for each dataset
SDR_N_All   <- sum(  ALL_1_daily_DESEAS[,GLB_att_N])
SDR_N_Clear <- sum(CLEAR_1_daily_DESEAS[,GLB_att_N])
SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N])
## TSI data span info
TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
                    strip.white = TRUE,
                    sep = ";",
                    header = TRUE))
## variables we use in the analysis
source("./DHI_GHI_0_variables.R")

## create local citations file and links here
source("~/CODE/FUNCTIONS/R/gather_citations.R")
suppressMessages(
    gather_citations(
        sourcefile   = "./Article.Rmd",
        mainreposit  = "./Mainreferences.bib",
        localreposit = "./references.bib",
        folderlink   = "./References_links",
        workingdir   = "./"
    )
)
```


# Introduction.

<!--
Η εισαγωγή είναι εντάξει. Ίσως θα μπορούσες να βάλεις στη 2η
παράγραφο ακόμη 2-3 σειρές για τις τάσεις του αεροζόλ και των νεφών
από τις εργασίες και να καταλήξεις με κάποια στοιχεία για το αεροζόλ
στη Θεσσαλονίκη από εργασίες δικές μας. Για τα νέφη δεν έχω υπόψη μου
καμία εργασία αλλά ίσως θα μπορούσε να ειπωθεί κάτι χρησιμοποιώντας
παγκόσμιες εκτιμήσεις για την περιοχή της Μεσογείου.
π.χ. https://link.springer.com/article/10.1007/s00382-022-06438-3
-->


The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
Can also affect, solar and agricultural production, and all living organisms.
Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
<!-- Trends china -->
On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
<!-- Trends Europe Brazil -->
A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
<!-- Trends Global -->
Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend [@Yuan2021].

There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds. 
Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
Similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
<!-- Sun variability: should cite that there is no significant trend found Wild2021 -->

Due to the variability of the phenomenon, and its contributing factors, there is a constant need to investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
<!-- goals -->
We reevaluated and extended the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derive the radiation trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).


<!--
Aerosols interactions

#Yang2021  Cloud “shrinking” and “optical thinning” in the “dimming” period and a subsequent recovery in the “brightening” period over China

Aerosols not only strengthen but also weaken the
growth of clouds with respect to their coverage and
optical thickness, depending on the levels of pollution
and the associated amounts of aerosols as postulated
in a conceptual framework (Wild 2009a, 2012, Yang
et al 2012). 


...... cloud formation occurrences and constitution nuclei. cite .....

#Wild2012 : simplify interaction nice graph

-->

# Observational data and methodology.

The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki 
($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.)
in the period from 
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS$Date)`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS$Date)`.
<!-- }. -->
The monitoring site is located near the city centre, and we expect to be affected by the urban environment. 
During the study period, the pyranometer has been independently calibrated three times at the Meteorologisches Observatorium Lindenberg, DWD, when it was verified the stability of the instrument to within better than $0.7\%$ relative to the initial calibration by the manufacturer. 
Along with SDR, the direct beam radiation (DNI) was also measured by a collocated Kipp & Zonen CHP-1 pyrheliometer, for the period
<!-- \mbox{ -->
`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`
<!-- } -->
to
<!-- \mbox{ -->
`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
<!-- }. -->
Although, we have performed a similar analysis to the DNI data the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data. 
However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
It is noted that despite the capability of the CSid algorithm to use the DNI as a characterization parameter, we haven't utilized it here, to avoid any selection bias, due to unequal length of the two datasets. 
There are four distinct steps in the creation of the dataset analysed here:
a)\ the acquisition of radiation measurements from the sensors,
b)\ the data quality check,
c)\ the identification of "clear sky" conditions from the radiometric data, and 
d)\ the aggregation of data and trend analysis.


<!--
The horizon is clear of obstructions except for the azimuth angles between 35 and 120 where obstacles of up to 10 elevation block the direct irradiance in the morning for SZA larger than 80.  These data points were excluded from the analysis, as
-->

<!-- Zerefos 
Northern Greece has been selected because it is well known to be located at the crossroads of aerosols and various pollutants pathways originating in central and Eastern
Europe (Lelieveld et al., 2002; Zerefos et al., 2002). Thessa-
loniki, the major city in northern Greece, lies along the exit of air
pollutants from Europe, in the path of the prevailing long-range
transport (LRT) trajectories, as we move from central and east-
ern Europe to the Mediterranean (Zerefos et. al., 2002). On the
-->

<!-- a) acquisition -->
For the acquisition of radiometric data, the signal of the pyranometer is sampled with a rate of $1\,\text{Hz}$.
The mean and the standard deviation of these samples are recorded every minute.
The measurements are corrected for the zero offset ("dark signal" in volts).
The "dark signal" is calculated by averaging all measurements recorded for a period of 
$3\,\text{h}$, 
before (morning) or after (evening) the Sun reaches an elevation angle of
$-10^\circ$.
The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between calibrations.

<!-- b) Radiation data quality check -->
A manual screening was performed, to remove inconsistent and erroneous recordings that can occur stochastically or systematically, during the continuous operation of the instruments.
The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, which is based on the methods of 
Long and Shi\ [-@Long2008a; -@Long2006].
Thus, problematic recordings have been excluded from further processing. 
Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiometric measurements used in this study.

<!-- c) Clear sky identification -->
In order to be able to estimate the effect of the sky condition on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
To identify the clear-sky conditions we used a method proposed by
@Long2000
and by
@Reno2016,
which was adapted and configured for the site, as the authors suggest.

<!-- CS subjectivity -->
We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in any method of characterization.
As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky dome.
Although the fine-tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.


```{r setup, echo=F, include=T, eval=TRUE}
## Load Clear Sky id v14.2 variables
MS <- data.frame(readRDS("~/CS_id/PARAMS/Clear_sky_id_Reno-Hansen_apply_v14_2.Rds"))
# export all Clear Sky variables
for (an in names(MS)) {
assign(an, MS[[an]])
}
## get TSI slope for reference
TSI <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
           strip.white = TRUE,
           sep = ";",
           header = TRUE))
TSI <- TSI[grep("TSI", TSI$var), ]
TSI$slope.sd           <- NULL
TSI$slope.ConfInt_0.95 <- NULL
TSI$slope.ConfInt_0.99 <- NULL
TSI$var                <- NULL
TSI$N                  <- NULL
TSI$Obs                <- NULL
setorder(TSI, -slope.stat_sig)
rownames(TSI) <- NULL
names(TSI) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
TSIslopeAll <- TSI[`Sky condition` == "All sky cond."]
```

<!-- # Detection of clear periods in GHI measurements for SDR trends. -->

For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
To calculate the reference clear sky
$\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by
the radiation model of @Haurwitz1945, adjusted for our site with a
factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described
by @Long2000 and @Reno2016.
The target of
the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023]. 
The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
These models are described in more details by @Reno2012 and evaluated by @Reno2016.
We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), 
<!--[....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}-->
while the procedure, manages to characterize the majority of the data.
<!--[]{.comment-end id="3"}--> 
Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.

<!-- cost fuction -->
\begin{equation}
f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
\end{equation}
where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.


<!-- adjusted Haurwitz -->
\begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = `r signif(alpha,3)` \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
\end{equation}
where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\,\text{m}^{-2}$ and $\theta$ is the solar zenith angle (SZA).


The criteria that were used to identify whether a measurement was taken under clear-sky conditions are presented below.
A data point is flagged as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
Each criterion was applied for a running window of $`r nt`$ consecutive one-minute measurements, and the characterization is assigned to the central value of the window.
Each parameter, was calculated both from the observations and the reference clear sky model, for each comparison.
The allowable range of variation is defined by the model-derived value of the parameter multiplied by a factor plus an offset. 
The factors and the offsets were determined empirically, by manual inspecting each filters performance on selected days, and adjusting them accordingly, during an iterative process.

<!-- 1. MeanVIP -->
<!-- ### Mean value of irradiance during the time period. -->

<!--a) Mean of the measured $\overline{\text{SDR}}_i$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$ and running mean of the reference $\overline{\text{SDR}}_{\text{CSref},i}$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq. \ref{eq:MeanVIP}).-->

a) Mean of the measured $\overline{\text{SDR}}_i$ (Eq. \ref{eq:MeanVIP}).
\begin{equation}
`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{\text{CSref},i} - `r CS_ref_rm_VIP_LowOff`
< \overline{\text{SDR}}_i <
`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{\text{CSref},i} + `r CS_ref_rm_VIP_UppOff`
\label{eq:MeanVIP}
\end{equation}


<!-- 2. MaxVIP -->
<!-- ### Max value of irradiance during the time period. -->

<!-- The running max measured value $M_{\text{SDR},i} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$, is compared to a similar constructed value from the reference $M_{CSrefi} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.\ \ref{eq:MaxVIP}). -->

b) Maximum measured value $M_{\text{}}$ (Eq.\ \ref{eq:MaxVIP}).
\begin{equation}
`r MaxVIP_fct` \times M_{\text{CSref},i} - `r MaxVIP_off_low`
< M_{\text{}i} <
`r MaxVIP_fct` \times M_{\text{CSref},i} + `r MaxVIP_off_upp`
\label{eq:MaxVIP}
\end{equation}

<!-- 3. VIL -->
<!-- ### Variability in irradiance by line length. -->
c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Eq. \ref{eq:VILeq}).
\begin{equation}
L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
\label{eq:VILeq}
\end{equation}
\begin{equation}
`r MinVIL_fct` \times L_{\text{CSref},i} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` \times L_{\text{CSref},i} + `r offVIL_upl`
\label{eq:VILcr}
\end{equation}
where: $t_i$ is the time each SDR measurement has been measured

<!-- 4. VCT -->
<!-- ### Variance of Changes in the Time series. -->

<!-- Standard deviation of rate of change in irradiance. -->
d) Standard deviation $\sigma_i$ of the slope ($s_i$) between the $`r nt`$ sequential points, normalized by the mean $\overline{\text{SDR}}_i$ (Eq.\ \ref{eq:VCT1}).
<!-- \begin{gather}
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
\end{gather} -->
\begin{gather}
  \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
  s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
\end{gather}
For this criterion, $\sigma_i$ should be below a certain threshold (Eq.\ \ref{eq:VCTcr}):
\begin{equation}
  \sigma_i < `r offVCT` \label{eq:VCTcr}
\end{equation}

<!-- 5. VSM -->
<!-- ### Variability in the Shape of the irradiance Measurements. -->

e) Maximum difference $X_i$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
\begin{gather}
  X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
  x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
  x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
\end{gather}
For this criterion, $X_i$ should be below a certain threshold (Eq.\ \ref{eq:VSMcr}):
\begin{equation}
  X_i < `r offVSM` \label{eq:VSMcr}
\end{equation}


<!-- ### Data selection -->
Due to a significant measurement uncertainty near the horizon, we have to exclude all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between 
$35^\circ$ and $120^\circ$ with SZA greater than $80^\circ$.
On the latter instances, Sun is systematically, not visible by the instrument's location.
<!-- 1au -->
To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
<!-- TSI -->
Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a <!--homogenized--> time series of satellite TSI observations.
The TSI data we use are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005]. 
Where the initial daily values, were interpolated to match with the time step of our measurements. 
The final dataset contains
$`r SDR_N_All`$
one-minute measurements, of which, 
$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
were identified as under clear-sky conditions and subsequently
$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
as under cloud-sky conditions.
<!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->

<!-- aggregation -->
In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser timescale.
To preserve the representativeness of the data we used the following criteria: 
<!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
a) for the daily mean values we accept days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements, present and valid,
b) monthly values were computed from daily means
only when at least `r Monthly_aggegation_N_lim` days were available.
To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
<!-- Seasonal -->
For the seasonal means we averaged the mean daily values in each season (Winter: December - February, Spring: March - May, etc.).
<!-- Deseasonalization -->
Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
To estimate SZA contribution to the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.



# Results

## Long-term SDR trends






# Version

This Rmd-skeleton uses the mdpi Latex template published 2023-03-25. 
However, the official template gets more frequently updated than the **rticles**
package. Therefore, please make sure prior to paper submission, that you're 
using the most recent .cls, .tex and .bst files 
(available [here](http://www.mdpi.com/authors/latex)).

# Article Header Information




## Journal Specific YAML variables






# Introduction

The introduction should briefly place the study in a broad context and highlight
why it is important. It should define the purpose of the work and its
significance. The current state of the research field should be reviewed
carefully and key publications cited. Please highlight controversial and
diverging hypotheses when necessary. Finally, briefly mention the main aim of
the work and highlight the principal conclusions. As far as possible, please
keep the introduction comprehensible to scientists outside your particular
field of research. Citing a journal paper [@bertrand-krajewski_distribution_1998;
@leutnant_stormwater_2016]. And now citing a book reference @gujer_systems_2008.
Some MDPI journals use Chicago and others use APA, this template should choose
the correct citation format for you once you specify the journal in the YAML
header.

To use endnotes, change `endnotes: true` in the YAML header, then use
`\endnote{This is an endnote.}`.


# Materials and Methods

Materials and Methods should be described with sufficient details to allow 
others to replicate and build on published results. Please note that publication 
of your manuscript implicates that you must make all materials, data, computer 
code, and protocols associated with the publication available to readers. Please
disclose at the submission stage any restrictions on the availability of 
materials or information. New methods and protocols should be described in 
detail while well-established methods can be briefly described and appropriately
cited.

Research manuscripts reporting large datasets that are deposited in a publicly 
available database should specify where the data have been deposited and provide
the relevant accession numbers. If the accession numbers have not yet been
obtained at the time of submission, please state that they will be provided
during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require
ethical approval must list the authority that provided approval and the
corresponding ethical approval code.

# Results

This section may be divided by subheadings. It should provide a concise and
precise description of the experimental results, their interpretation as well
as the experimental conclusions that can be drawn.

## Subsection Heading Here

Subsection text here.

### Subsubsection Heading Here

Bulleted lists look like this:

* First bullet
* Second bullet
* Third bullet

Numbered lists can be added as follows:

1. First item
2. Second item
3. Third item

The text continues here.

## Figures, Tables and Schemes

All figures and tables should be cited in the main text as Figure \ref{fig:fig1}, 
\ref{tab:tab1}, etc. To get cross-reference to figure generated by R chunks 
include the `\\label{}` tag in the `fig.cap` attribute of the R chunk: 
`fig.cap = "Fancy Caption\\label{fig:plot}"`.

```{r fig1, echo=FALSE, fig.width=3, fig.cap="A figure added with a code chunk.\\label{fig:fig1}"}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```


When making tables using `kable`, it is suggested to use
the `format="latex"` and `tabl.envir="table"` arguments
to ensure table numbering and compatibility with the mdpi
document class.

```{r tab1, echo=FALSE}
knitr::kable(mtcars[1:5, 1:3], format = "latex", 
             booktabs = TRUE, 
             caption = "This is a table caption. Tables should be placed in the 
             main text near to the first time they are~cited.", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```



For a very wide table, landscape layouts are allowed.


\startlandscape

```{r tab2, echo=FALSE}
df <- data.frame(`Title 1` = c("Entry 1", "Entry 2"),
                 `Title 2` = c("Data", "Data"),
                 `Title 3` = c("Data", "Data"),
                 `Title 4` = c("This cell has some longer content that runs over
                               two lines",
                               "Data"))
knitr::kable(df, format = "latex", 
             booktabs = TRUE, 
             caption = "This is a very wide table", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```

\finishlandscape

## Formatting of Mathematical Components

This is an example of an equation:

$$
a = 1.
$$

If you want numbered equations use Latex and wrap in the equation environment:

\begin{equation}
a = 1,
\end{equation}

the text following an equation need not be a new paragraph. Please punctuate 
equations as regular text.

```{comment}
If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
```

This is the example 2 of equation:

\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

Theorem-type environments (including propositions, lemmas, corollaries etc.) 
can be formatted as follows:

Example of a theorem:

::: {.Theorem latex=true}
Example text of a theorem
:::



The text continues here.

# Discussion

Authors should discuss the results and how they can be interpreted in 
perspective of previous studies and of the working hypotheses. The findings and 
their implications should be discussed in the broadest context possible. Future 
research directions may also be highlighted.

# Conclusion

This section is not mandatory, but can be added to the manuscript if the
discussion is unusually long or complex.


